The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-02 paper.tex
resolving the artifact- and human-centered section lenght conflict

\paragraph{Artifact-centered perspective on anthropomorphism}

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize it. 
This perspective assumes that humans directly (mindlessly) respond to life-like and social 
cues that the non-human agent or system emits \citep{nass_machines_2000}. Without thoughtful mental processing, 
humans tend to simply apply stereotypes and heuristics to the non-human agent, and in turn apply human-human social schemas and norms to the occurring interactions.
\cite{takayama_perspectives_2012} propose to refer to this immediate (sometimes visceral) sense in a situation as \textit{in-the-moment}, compared to a more distanced cogitation and consideration which she calls \textit{reflective}. As this distinction implies different cognitive processes, we will come back to it later, in section \ref{sec:cognitive-expl}.

@@

In contrast to the artifact-centered perspective, in the human-centered perspective the argumentation is, that the human is motivated to anthropomorphize the non-human artifact). As already mentioned, both these views go hand in hand.

Humans tend to perceive human-like characteristics such as physical appearance, emotional states, or inner mental states and motivations in non-human agents that do not have these characteristics themselves, such as animals, natural forces, religious agents, technological gadgets, or mechanical devices \citep{epley_when_2008}.\footnote{According to \cite{epley_when_2008}, a non-human agent can be anything that acts -- or is believed to act -- with apparent independence.}  Anthropomorphism has received remarkable attention across various different domains, such as natural sciences, developmental and social psychology, Human-Computer Interaction (HCI), and Human-Robot Interaction (HRI).

@@

\subsubsection{Why do people anthropomorphize?\\}
\label{sec:explanations}

But why do people anthropomorphize? First and foremost, it needs to be mentioned that anthropomorphism is a fairly complex phenomenon. Paradoxically, there is a huge and diverse body of literature on it but nevertheless, it seems that anthropomorphism is still not very well understood (especially in a field as young as HRI).

One common reason for anthropomorphism is that by perceiving some human-like characteristics in the non-human agent, this agent is made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one can feel \emph{more empathetic} toward it, and possibly one can then anticipate the agent's behavior. These are central motivations when trying to explain anthropomorphism, and we will come back to them in more detail later on. 
It is described that for the just mentioned reasons, people commonly anthropomorphize their pets, by ascribing mental states, intentions or feelings to them\footnote{Interestingly, pet-ownership seems to have a significant impact on human-robot interaction: several studies suggest that pet-owners are more likely to anthropomorphize technologies and robots than non pet-owners} \cite{eddy_attribution_1993}. But how can these motivations behind anthropomorphism be explained from a human-centered point of view?

@@

This perspective builds on research from developmental and social psychology as well as cognitive sciences.  In the following, we present what we found in reviewing this literature: (1) a line of research in developmental psychology that studies the attribution of animacy; (2) a line of research in more cognitive sciences that studies the cognitive processes that underlie anthropomorphism, and (3) a theory that explains anthropomorphism based on psychological factors~\citep{epley_seeing_2007}. All three branches apply a human-centered perspective on anthropomorphism.

>>>>

\paragraph{Artifact-centered perspective on anthropomorphism\\}

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize it. 
This perspective assumes that humans directly (mindlessly) respond to life-like and social 
cues that the non-human agent or system emits \citep{nass_machines_2000}. Without thoughtful mental processing, 
humans tend to simply apply stereotypes and heuristics to the non-human agent, and in turn apply human-human social schemas and norms to the occurring interactions.
\cite{takayama_perspectives_2012} propose to refer to this immediate (sometimes visceral) sense in a situation as \textit{in-the-moment}, compared to a more distanced cogitation and consideration which she calls \textit{reflective}. As this distinction implies different cognitive processes, it is more approriate to apply a human-centered perspective and we will come back to it later, in section \ref{sec:cognitive-expl}.

@@

\paragraph{Human-centered perspective on anthropomorphism}

In contrast to the artifact-centered perspective, in the human-centered perspective the argumentation is, that the human is motivated to anthropomorphize the non-human artifact. As already mentioned, both these views go hand in hand.

Humans tend to perceive human-like characteristics such as physical appearance, emotional states, or inner mental states and motivations in non-human agents that do not have these characteristics themselves, such as animals, natural forces, religious agents, technological gadgets, or mechanical devices \citep{epley_when_2008}.\footnote{According to \cite{epley_when_2008}, a non-human agent can be anything that acts -- or is believed to act -- with apparent independence.} 
One common reason for anthropomorphism is that by perceiving some human-like characteristics in the non-human agent, this agent is made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one can feel \emph{more empathetic} toward it, and possibly one can then anticipate the agent's behavior. These are central motivations when trying to explain anthropomorphism, and we will come back to them in more detail later on. 
It is described that for the just mentioned reasons, people commonly anthropomorphize their pets, by ascribing mental states, intentions or feelings to them\footnote{Interestingly, pet-ownership seems to have a significant impact on human-robot interaction: several studies suggest that pet-owners are more likely to anthropomorphize technologies and robots than non pet-owners} \cite{eddy_attribution_1993}. But how can these motivations behind anthropomorphism be explained from a human-centered point of view?

In the following, we present what we found in reviewing literature that applies a human-centered perspective on anthropomorphism by building on:
\textit{(1)} a line of research in developmental psychology that studies the attribution of animacy;
\textit{(2)} a line of research in more cognitive sciences that studies the cognitive processes that underlie anthropomorphism, and
\textit{(3)} a theory that explains anthropomorphism based on psychological factors.

@@

%\subsubsection{Why do people anthropomorphize?\\}
%\label{sec:explanations}

%But why do people anthropomorphize? First and foremost, it needs to be mentioned that anthropomorphism is a fairly complex phenomenon. Paradoxically, there is a huge and diverse body of literature on it but nevertheless, it seems that anthropomorphism is still not very well understood (especially in a field as young as HRI).

@@
